---
title: "27-COUNTY-OVERALL-youth-missing-graphics-calculations"
output: html_notebook
---

The purpose of this notebook is to create a data frame that we can use to plot the Overall youth missing graphic from Figures 16-22 on Pages 115-118 of the Access Denied Report.We will call the final data frame "Overall_cn_df."

# Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,tidyr,vroom,stringr, readr, assertr, ggplot2)
```

Now we can use the original wide data to calculate the missing percentages for different demographic groups and then tack those calculations on to the original data frame "cn_and_state_overall_df" that we created in the 27-COUNTY-youth-missing-graphics-calculations RMD file.

### Overall

We will start by finding percent missingness in the total population


Now we need to break down to populations by students who are in schools that offer GT programs and schools that do not offer GT programs

```{r}
# Find the total number of students in schools that do not offer GT services
ct_enrollment_no_gt_df <- ocr_17_wide %>%
  filter(SCH_GT_IND == "No") %>%
  group_by(LEA_STATE, NMCNTY) %>%
  summarise(ct_tot_enr_no_gt = (sum(TOT_ENR_M,na.rm = TRUE) + sum(TOT_ENR_F,na.rm = TRUE))) %>%
  ungroup()
  
```

```{r}
# Find the total number of students in schools that offer GT services
ct_enrollment_w_gt_df <- ocr_17_wide %>%
  filter(SCH_GT_IND == "Yes") %>%
  group_by(LEA_STATE, NMCNTY) %>%
  summarise(ct_tot_enr_w_gt = (sum(TOT_ENR_M,na.rm = TRUE) + sum(TOT_ENR_F,na.rm = TRUE))) %>%
  ungroup()
```

```{r}
# Add number of students from schools with different service types to full DF
OV_ct_missing_df <- cn_and_state_overall_df %>%
  left_join(ct_enrollment_no_gt_df, by = c("LEA_STATE" = "LEA_STATE", "NMCNTY" = "NMCNTY")) %>%
  left_join(ct_enrollment_w_gt_df, by = c("LEA_STATE" = "LEA_STATE", "NMCNTY" = "NMCNTY"))

```

# Clean NAs out of columns

```{r}
# Clean NAs from the columns for gt and non gt student enrollment
OV_ct_missing_df[,12:13][is.na(OV_ct_missing_df[,12:13])] <- 0
```

# Calculate Missingness
Now our data frame has all the columns we need to calculate the number of missing AM students from gt programs.

```{r}
# Calculate missingness counts
OV_ct_missing_df <- OV_ct_missing_df %>%
    mutate(
          OV_no_gt_missing_ct_lower = case_when(is.na(ct_perc_gt) == TRUE ~ NA_real_,
                                             state_perc_gt > 0 ~ round(state_perc_gt * ct_tot_enr_no_gt), 
                                             state_perc_gt == 0 ~ round(national_gt_avg * ct_tot_enr_no_gt),
                                             ), 
          OV_no_gt_missing_ct_upper = case_when(tot_ct_enr == 0 ~ NA_real_,
                                             state_perc_gt > 0 ~ round(state_nonT1_perc_gt * ct_tot_enr_no_gt), 
                                             state_perc_gt == 0 ~ round(national_gt_avg_nont1 * ct_tot_enr_no_gt),
                                             ), 
          OV_underrep_missing_ct_lower = case_when(tot_ct_enr == 0 ~ NA_real_,
                                                    tot_ct_enr > 0 & ct_tot_enr_w_gt == 0 ~ 0,
                                                    tot_ct_enr > 0 & ct_tot_enr_w_gt > 0 ~ round((ct_tot_enr_w_gt *                                                     state_perc_gt)) - tot_ct_gt_enr
                                             ),
          OV_underrep_missing_ct_upper = case_when(tot_ct_enr == 0 ~ NA_real_,
                                                    tot_ct_enr > 0 & ct_tot_enr_w_gt == 0 ~ 0,
                                                    tot_ct_enr > 0 & ct_tot_enr_w_gt > 0 ~ round((ct_tot_enr_w_gt *                                                     state_nonT1_perc_gt)) - tot_ct_gt_enr
                                             )
         )
```

```{r}
#Convert negative underrep counts to 0
OV_ct_missing_df[,16:17][(OV_ct_missing_df[,16:17]) < 0] <- 0
```

```{r}
# Generate counts and percentages of missing students
OV_ct_missing_df <- OV_ct_missing_df %>%
  mutate(
         OV_count_missing_ct_lower = OV_no_gt_missing_ct_lower + OV_underrep_missing_ct_lower,
         OV_count_missing_ct_upper = OV_no_gt_missing_ct_upper + OV_underrep_missing_ct_upper,
         OV_perc_missing_ct_lower = (OV_no_gt_missing_ct_lower + OV_underrep_missing_ct_lower)/(OV_no_gt_missing_ct_lower +                                               OV_underrep_missing_ct_lower + tot_ct_gt_enr),
         OV_perc_missing_ct_upper = (OV_no_gt_missing_ct_upper + OV_underrep_missing_ct_upper)/(OV_no_gt_missing_ct_upper +                                               OV_underrep_missing_ct_upper + tot_ct_gt_enr)
         )
```

Now we have all the missing percentages calculated we need to do a bit of cleaning before we can plot this data

```{r}
# Select only the columns we want to plot information from
OV_ct_df <- OV_ct_missing_df %>%
  select(LEA_STATE, NMCNTY, OV_count_missing_ct_lower, OV_count_missing_ct_upper, OV_perc_missing_ct_lower, OV_perc_missing_ct_upper)
```

```{r}
#Save the full data frame to a CSV file
write_csv(OV_ct_df, "C:/Users/donse/OneDrive/Documents/DSSG/gateways/OV_ct_df.csv")
```

