---
title: "10-load-data"
output: html_notebook
---

The purpose of this notebook is to load and clean the data.  It should also test the data (e.g., by assertr) to ensure that assumptions about the data are met, and store the cleaned and tested data.

# Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,tidyr,vroom,stringr,arrow,testit)
source(knitr::purl("12-subpopulation_function.Rmd", output = tempfile()))
source(knitr::purl("13-enrollment-fixing-functions-for-long-data.Rmd", output = tempfile()))
source(knitr::purl("14-enrollment-fixing-functions-for-wide-data.Rmd", output = tempfile()))
```


# Add LOCALE data from CCD
## Define function
```{r}
add_locale_cnty_title <- function(ocr_df, ccd_locale, ccd_title, ocr_ccd_crosswalk, begin_year = '2015'){
  
  # Remove COMBOKEY without match from crosswalk df
  no_match_ii = which(ocr_ccd_crosswalk$Match ==FALSE & ocr_ccd_crosswalk$`Additional Match` == FALSE)
  ocr_ccd_crosswalk = ocr_ccd_crosswalk[-no_match_ii,]
  
  # Select only relevant columns from CCD dfs
  if(begin_year == '2015'){
    
    ccd_locale = ccd_locale %>% 
      select(NCESSCH, LOCALE15, CNTY15, NMCNTY15)
    ccd_title = ccd_title %>% 
      select(NCESSCH, TITLEI)
    
  }else if(begin_year == '2017'){
    
    ccd_locale = ccd_locale %>% 
      select(NCESSCH, LOCALE, CNTY, NMCNTY)
    ccd_title = ccd_title %>% 
      select(NCESSCH, TITLEI_STATUS_TEXT)
    
  }
  
  # Select only relevant columns from crosswalk df (year of interest)
  ocr_ccd_crosswalk = ocr_ccd_crosswalk[,c(1,grep(begin_year, colnames(ocr_ccd_crosswalk)))] %>%
    tidyr::drop_na()
  
  # Inner join to merge CCD and Crosswalk dfs
  ccd_locale_crosswalk = inner_join(ccd_locale, ocr_ccd_crosswalk, by = c("NCESSCH" = colnames(ocr_ccd_crosswalk)[2]))
  ccd_locale_crosswalk['NCESSCH'] = NULL
  ccd_title_crosswalk = inner_join(ccd_title, ocr_ccd_crosswalk, by = c("NCESSCH" = colnames(ocr_ccd_crosswalk)[2]))
  ccd_title_crosswalk['NCESSCH'] = NULL
  
  # Left join to add LOCALE and TITLE to OCR data
  ocr_locale_title = left_join(ocr_df, ccd_locale_crosswalk, by = "COMBOKEY")
  ocr_locale_title = left_join(ocr_locale_title, ccd_title_crosswalk, by = "COMBOKEY")
  
  ocr_locale_title
}
```

# Load data & select columns of interest

```{r}
setClass(Class = "OCR",
         representation(
             year = "numeric",
             key = "data.frame",
             wide = "data.frame",
             long = "data.frame"             
         ))

```

```{r}

transOCR2017 <- function(){
    #"FUNCTION DESCRIPTION HERE"

    ## Open record layout documents and select variable names for identifying schools,
    ## enrollment, and gifted & talented students.
    ocr_key1 <- readxl::read_xlsx("data/OCR/2017-18-crdc-data/2017-18-crdc-data-corrected-publication 2/2017-18 Public-Use Files/Documentation/2017-18 CRDC File Structure.xlsx",
                                 sheet = "Gifted and Talented")

    ocr_key2 <- readxl::read_xlsx("data/OCR/2017-18-crdc-data/2017-18-crdc-data-corrected-publication 2/2017-18 Public-Use Files/Documentation/2017-18 CRDC File Structure.xlsx",
                                  sheet = "Enrollment") %>%
        filter(grepl("SCH_ENR|TOT_ENR",`Field Name` ))

    ocr_key <- rbind(ocr_key1, ocr_key2)

    ## Open full school-level data files, import only variables of interest, convert large numeric vars to chars, merge
    df_ocr_ENR_wide <- vroom("data/OCR/2017-18-crdc-data/2017-18-crdc-data-corrected-publication 2/2017-18 Public-Use Files/Data/SCH/CRDC/CSV/Enrollment.csv",
                         col_select = c(ocr_key2$`Field Name`, 'COMBOKEY'),
                         col_types = c(COMBOKEY = "c"))
    df_ocr_GATE_wide <- vroom("data/OCR/2017-18-crdc-data/2017-18-crdc-data-corrected-publication 2/2017-18 Public-Use Files/Data/SCH/CRDC/CSV/Gifted and Talented.csv",
                         col_select = c(ocr_key1$`Field Name`, 'COMBOKEY', 'LEAID', 'SCHID'),
                         col_types = c(COMBOKEY = "c",
                                       LEAID = "c",
                                       SCHID = "c"))
    df_ocr_wide <- merge(df_ocr_GATE_wide, df_ocr_ENR_wide, by = 'COMBOKEY')
    
    ## Build lookup table
    ocr_lookup <- unique(df_ocr_wide[,c("LEA_STATE","LEAID","LEA_NAME","SCHID","SCH_NAME","COMBOKEY")])
    df_ocr_wide <- df_ocr_wide %>% select(-c(LEA_NAME,
                                             #LEAID,
                                             #LEA_STATE,
                                             #SCH_GT_IND,
                                             LEA_STATE_NAME,
                                             #SCHID,
                                             SCH_NAME,
                                             JJ))
    
    # Load CCD data & crosswalk file
    ## WARNING: CCD Title I data for 2017-2018 is a version behind the 2015-2016 data & does not contain the same
    ## variable names or data
    ccd_locale <- readxl::read_xlsx('data/CCD/2017-2018/EDGE_GEOCODE_PUBLICSCH_1718/EDGE_GEOCODE_PUBLICSCH_1718.xlsx',
                                    col_types = c(NCESSCH = 'text'))
    ccd_title <- vroom("data/CCD/2017-2018/ccd_sch_129_1718_w_1a_083118.csv",
                          col_types = c(NCESSCH = 'c'))
    ocr_ccd_crosswalk <- readxl::read_xlsx("data/2017-18-CRDC-EDFacts-Crosswalk-rev.xlsx",
                                           sheet = "2017-18 Crosswalk",
                                           col_types = c(COMBOKEY = 'text'))
    
    ## Add locale data
    
    df_ocr_wide <- add_locale_cnty_title(ocr_df = df_ocr_wide, 
                                ccd_locale = ccd_locale, 
                                ccd_title = ccd_title, 
                                ocr_ccd_crosswalk = ocr_ccd_crosswalk, 
                                begin_year = '2017')
    
    #TITLEI
    df_ocr_wide['TITLEI_original'] <- df_ocr_wide$TITLEI_STATUS_TEXT
    ii <- which(df_ocr_wide$TITLEI_STATUS_TEXT %in% c("Missing", "Not reported"))
    df_ocr_wide$TITLEI_STATUS_TEXT[ii] <- "Unknown"
    ii <- which(df_ocr_wide$TITLEI_STATUS_TEXT %in% c("Not a Title I school"))
    df_ocr_wide$TITLEI_STATUS_TEXT[ii] <- "No"
    df_ocr_wide$TITLEI_STATUS_TEXT[-ii] <- "Yes"
    
    # Gifted ID column
    ii <- which(df_ocr_wide$SCH_GT_IND %in% c("Yes", "No"))
    df_ocr_wide$SCH_GT_IND[-ii] <- "No" # to match Anne's numbers
    
    # LOCALE column
    df_ocr_wide['LOCALE_original'] <- df_ocr_wide$LOCALE
    city_ii <- which(df_ocr_wide$LOCALE %in% c('11', '12', '13'))
    suburb_ii <- which(df_ocr_wide$LOCALE %in% c('21', '22', '23'))
    town_ii <- which(df_ocr_wide$LOCALE %in% c('31', '32', '33'))
    rural_ii <- which(df_ocr_wide$LOCALE %in% c('41', '42', '43'))
    df_ocr_wide$LOCALE[-c(city_ii, suburb_ii, town_ii, rural_ii)] <- "Unknown"
    df_ocr_wide$LOCALE[city_ii] <- "City"
    df_ocr_wide$LOCALE[suburb_ii] <- "Suburb"
    df_ocr_wide$LOCALE[town_ii] <- "Town"
    df_ocr_wide$LOCALE[rural_ii] <- "Rural"
    
    # Standardize new column names
    colnames(df_ocr_wide)[grep('^TITLEI_STATUS_TEXT$', colnames(df_ocr_wide))] <- 'TITLEI'
    
    ## Transform wide dataset to long
    df_ocr_long <- df_ocr_wide %>% 
        gather(module, enrollment, SCH_GTENR_HI_M:SCH_ENR_IDEA_F, factor_key = T)

    df_ocr_long <- df_ocr_long %>% mutate(gender = str_extract(module, "[FM]$"),
                                          race = str_extract(module, "(HI)|(MHI)|(AS)|(AM)|(WH)|(BL)|(TR)|(HP)"),
                                          other = str_extract(module, "(IDEA)(?=_)|(LEP)(?=_)|(504)(?=_)"),
                                          module = str_extract(module, "SCH_([A-Z]+|[0-9])+|TOT_([A-Z]+|[0-9])+"))
    
    ## Apply enrollment fixes for wide data
    df_ocr_wide <- enrollment_fix_wide(df_ocr_wide)
    ## Apply enrollment fixes for long data
    df_ocr_long <- enrollment_fix_long(df_ocr_long)
    
    ## Return OCR object
    return(new("OCR",
               year = 2017,
               key = ocr_key,
               wide = df_ocr_wide,
               long = df_ocr_long))
    
}

transOCR2015 <- function(){
    #"FUNCTION DESCRIPTION HERE"

    ## Open record layout document and select variable names for identifying schools,
    ## enrollment, and gifted & talented students.
    ocr_key <- vroom("data/OCR/2015-16-crdc-data/Data Files and Layouts/CRDC 2015-16 School Data Record Layout.csv")
    columns <- ocr_key$Field_Name[ocr_key$Module == "Identification" | 
                                  ocr_key$Module == "Enrollment" |
                                  ocr_key$Module == "Gifted and Talented"]

    ## Open full school-level data file, import only variables of interst, convert large numeric vars to char
    df_ocr_wide <- vroom("data/OCR/2015-16-crdc-data/Data Files and Layouts/CRDC 2015-16 School Data.csv", 
                         col_select = columns,
                         col_types = c(COMBOKEY = "c",
                                       LEAID = "c",
                                       SCHID = "c"))

    ## Build lookup table
    ocr_lookup <- unique(df_ocr_wide[,c("LEA_STATE","LEAID","LEA_NAME","SCHID","SCH_NAME","COMBOKEY")])
    df_ocr_wide <- df_ocr_wide %>% select(-c(LEA_NAME,
                                             #LEAID,
                                             #LEA_STATE,
                                             #SCH_GT_IND,
                                             LEA_STATE_NAME,
                                             #SCHID,
                                             SCH_NAME,
                                             JJ) & 
                                            !(contains("PSENR")))
    
    ### Helpful functions to pad SCHID and combine with LEAID to create new unique COMBOKEY
    df_ocr_wide <- df_ocr_wide %>%
      mutate(SCHID = str_pad(SCHID, 5, side='left', pad='0')) %>%
      mutate(LEAID = str_pad(LEAID, 7, side='left', pad='0')) 
    df_ocr_wide <- df_ocr_wide %>%
      mutate(COMBOKEY = str_c(LEAID, SCHID)) %>%
      select(COMBOKEY, everything())
    ### Assert to verify the functionality of the necessary recoding
    assert_12_characters <- function(x) nchar(x)==12
    assert(assert_12_characters(df_ocr_wide$COMBOKEY))

    ## Load CCD data & crosswalk file
    ccd_locale <- readxl::read_xlsx('data/CCD/2015-2016/EDGE_GEOCODE_PUBLICSCH_1516/EDGE_GEOCODE_PUBLICSCH_1516.xlsx', 
                                    col_types = c(NCESSCH = 'text'))
    ccd_title <- vroom("data/CCD/2015-2016/ccd_sch_129_1516_w_2a_011717.csv", 
                          col_types = c(NCESSCH = 'c'))
    ocr_ccd_crosswalk <- readxl::read_xlsx("data/2017-18-CRDC-EDFacts-Crosswalk-rev.xlsx", 
                                           sheet = "2017-18 Crosswalk", 
                                           col_types = c(COMBOKEY = 'text'))

    ## Add locale data
    
    df_ocr_wide <- add_locale_cnty_title(ocr_df = df_ocr_wide, 
                                ccd_locale = ccd_locale, 
                                ccd_title = ccd_title, 
                                ocr_ccd_crosswalk = ocr_ccd_crosswalk, 
                                begin_year = '2015')
    
    #TITLEI
    df_ocr_wide['TITLEI_original'] <- df_ocr_wide$TITLEI
    ii <- which(df_ocr_wide$TITLEI %in% c("Yes", "No"))
    df_ocr_wide$TITLEI[-ii] <- "Unknown"
    # Gifted ID column
    ii <- which(df_ocr_wide$SCH_GT_IND %in% c("Yes", "No"))
    df_ocr_wide$SCH_GT_IND[-ii] <- "No" # to match Anne's numbers
    # LOCALE15 column
    df_ocr_wide['LOCALE15_original'] <- df_ocr_wide$LOCALE15
    city_ii <- which(df_ocr_wide$LOCALE15 %in% c('11', '12', '13'))
    suburb_ii <- which(df_ocr_wide$LOCALE15 %in% c('21', '22', '23'))
    town_ii <- which(df_ocr_wide$LOCALE15 %in% c('31', '32', '33'))
    rural_ii <- which(df_ocr_wide$LOCALE15 %in% c('41', '42', '43'))
    df_ocr_wide$LOCALE15[-c(city_ii, suburb_ii, town_ii, rural_ii)] <- "Unknown"
    df_ocr_wide$LOCALE15[city_ii] <- "City"
    df_ocr_wide$LOCALE15[suburb_ii] <- "Suburb"
    df_ocr_wide$LOCALE15[town_ii] <- "Town"
    df_ocr_wide$LOCALE15[rural_ii] <- "Rural"
    
    # Standardize new column names
    colnames(df_ocr_wide)[grep('^LOCALE15$', colnames(df_ocr_wide))] <- 'LOCALE'
    colnames(df_ocr_wide)[grep('^CNTY15$', colnames(df_ocr_wide))] <- 'CNTY'
    colnames(df_ocr_wide)[grep('^NMCNTY15$', colnames(df_ocr_wide))] <- 'NMCNTY'
    colnames(df_ocr_wide)[grep('^LOCALE15_original$', colnames(df_ocr_wide))] <- 'LOCALE_original'

    ## Transform wide dataset to long
    df_ocr_long <- df_ocr_wide %>% 
        gather(module, enrollment, SCH_ENR_HI_M:SCH_GTENR_IDEA_F, factor_key = T)

    df_ocr_long <- df_ocr_long %>% mutate(gender = str_extract(module, "[FM]$"),
                                          race = str_extract(module, "(HI)|(MHI)|(AS)|(AM)|(WH)|(BL)|(TR)|(HP)"),
                                          other = str_extract(module, "(IDEA)(?=_)|(LEP)(?=_)|(504)(?=_)"),
                                          module = str_extract(module, "SCH_([A-Z]+|[0-9])+|TOT_([A-Z]+|[0-9])+"))
    ## Apply enrollment fixes for wide data
    df_ocr_wide <- enrollment_fix_wide(df_ocr_wide)
    ## Apply enrollment fixes for long data
    df_ocr_long <- enrollment_fix_long(df_ocr_long)

    ## Return OCR object
    return(new("OCR",
               year = 2015,
               key = ocr_key,
               wide = df_ocr_wide,
               long = df_ocr_long))

}


``` 

```{r}

writeOCR <- function(ocr){

    data_path <- str_c("data/OCR-transformed/",ocr@year,"/ocr_",ocr@year)
    
    if (file.exists(paste0("data/OCR-transformed/", ocr@year))){

        write_feather(ocr@key, paste0(data_path,"_lookup.feather"))
        write_feather(ocr@wide, paste0(data_path,"_wide.feather"))
        write_feather(ocr@long, paste0(data_path,"_long.feather"))
        
    } else {
        dir.create(paste0("data/OCR-transformed/", ocr@year))

        write_feather(ocr@key, paste0(data_path,"_lookup.feather"))
        write_feather(ocr@wide, paste0(data_path,"_wide.feather"))
        write_feather(ocr@long, paste0(data_path,"_long.feather"))
    }
}

```

## Write data to feather

```{r}
## Load and transform 2017-2018 OCR data
ocr_2017 <- transOCR2017()
writeOCR(ocr_2017)

## Load and transform 2015-2016 OCR data
ocr_2015 <- transOCR2015()
writeOCR(ocr_2015)

```

# Helper functions
```{r}
# Read in data
ocr_wide_2015 = read_feather(file='data/OCR-transformed/2015/ocr_2015_wide.feather')
ocr_wide_2017 = read_feather(file='data/OCR-transformed/2017/ocr_2017_wide.feather')
```

## 2015
### Without LOCALE
```{r}
# Nationwide
df_nation_nolocale_2015 <- subpopulation_gentable(ocr_wide_2015, groupby_colnames = c('TITLEI', 'SCH_GT_IND'))
# By state
df_states_nolocale_2015 <- subpopulation_gentable(ocr_wide_2015, groupby_colnames = c('LEA_STATE','TITLEI', 'SCH_GT_IND'))
```

### With LOCALE
```{r}
# Nationwide
df_nation_locale_2015 <- subpopulation_gentable(ocr_wide_2015, groupby_colnames = c('TITLEI', 'SCH_GT_IND', 'LOCALE'))
# By state
df_states_locale_2015 <- subpopulation_gentable(ocr_wide_2015, groupby_colnames = c('LEA_STATE','TITLEI', 'SCH_GT_IND', 'LOCALE'))
```

## 2017
### Without LOCALE
```{r}
# Nationwide
df_nation_nolocale_2017 <- subpopulation_gentable(ocr_wide_2017, groupby_colnames = c('TITLEI', 'SCH_GT_IND'))
# By state
df_states_nolocale_2017 <- subpopulation_gentable(ocr_wide_2017, groupby_colnames = c('LEA_STATE','TITLEI', 'SCH_GT_IND'))
```

### With LOCALE
```{r}
# Nationwide
df_nation_locale_2017 <- subpopulation_gentable(ocr_wide_2017, groupby_colnames = c('TITLEI', 'SCH_GT_IND', 'LOCALE'))
# By state
df_states_locale_2017 <- subpopulation_gentable(ocr_wide_2017, groupby_colnames = c('LEA_STATE','TITLEI', 'SCH_GT_IND', 'LOCALE'))
```



