---
title: "13-enrollment-fixing-functions-for-long-data"
output: html_notebook
---

The purpose of this notebook is to create functions to make the enrollment variable useable for the long data format. 

# Load required packages


```{r, purl = FALSE}
#Load required packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,tidyr,vroom,stringr, assertr)
```

# Function to look at whether or not gifted education is offered at the school level

This first function pulls the rows out of the long data frame where the value of the enrollment variable is either "Yes" or "No" and puts them into their own data frame.

```{r}
#Create a function to remove the non numeric enrollment values

get_services_offered_by_school_df <- function(df) {
 df <- df %>%
    filter(enrollment == "Yes" | enrollment == "No")
 return(df)
}
```

```{r, purl = FALSE}
#Run the function and create a new data frame
sevice_offered_df <- get_services_offered_by_school_df(df_ocr_long)
```

The new "service_offered_df" is a data frame where the enrollment variable indicates whether or not each school has gifted services offered. 

# Function to fix anomolies in the enrollment variable

The enrollment_fix_long() function is one that will make the enrollment variable usable in the long data frame. It will convert the column to numeric and it will remove data anomalies like negative numbers. 

```{r}
# Create a function to make the enrollment variable usable

enrollment_fix_long <- function(df) {
# make sure only numeric values are left in the enrollment variable
 df <- df %>%
    filter(enrollment != "Yes" | enrollment != "No")
# Convert enrollment variable to numeric
 df$enrollment <- as.numeric(df$enrollment)
# Convert negative enrollments to NAs
 df[,10][df[,10] < 0] <- NA
# Return data frame with zeroed out enrollments
 return(df)
}
```


```{r, purl = FALSE}
# Run the function on the long data to fix the enrollment variable
df_ocr_long_fixed <- enrollment_fix_long(df_ocr_long)
```

Now we have two new long data frames where the enrollment variable is usable!
