---
title: "27-2017-COUNTY-MEGATABLE-youth-missing-graphics-calculations"
output: html_notebook
---

The purpose of this notebook is to create a data frame that we can use to plot the youth missing graphics for all demographics at the district level in the Shiny app from Figures 16-22 on Pages 115-118 of the Access Denied Report.We will call the final data frame "district_missing_df"

# Change the column names so we can combine them into one data frame

```{r}
colnames(AM_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

```{r}
colnames(AS_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

```{r}
colnames(BL_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

```{r}
colnames(HI_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

```{r}
colnames(HP_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

```{r}
colnames(TR_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

```{r}
colnames(WH_ct_df) <- c("LEA_STATE", "NMCNTY", "total_ct_enr", "total_ct_gt_enr", "group_enr_ct", "group_gt_enr_ct", "group_missing_count_lower", "group_missing_count_upper", "group_missing_perc_lower", "group_missing_perc_upper", "group", "year", "region")
```

# Bind all groups together to get the 2015 data frame

```{r}
# Stack upper and lower data frames on top of each other to get our final data frame for plotting. 
county_missing_df <- rbind(AM_ct_df, AS_ct_df, BL_ct_df, HI_ct_df, HP_ct_df, TR_ct_df, WH_ct_df)
```

# Add state gt percentages to the data frame so that we can plot them as the slopes of the lines in scatterplots.
# Use state missing totals df generated in the state level folder

```{r}
# Create data frame for statewide gt percentages
perc_gt_df <- state_missing_totals_df %>%
  select(LEA_STATE, state_perc_gt, state_nonT1_perc_gt)
```


```{r}
# Add statewide gt percentages to district df
county_missing_df <- county_missing_df %>%
  left_join(perc_gt_df, by = c("LEA_STATE" = "LEA_STATE"))
```




