---
title: "27-AIAN-youth-missing-graphics-calculations"
output: html_notebook
---

The purpose of this notebook is to create a data frame that we can use to plot the AIAN youth missing graphic from Figures 16-22 on Pages 115-118 of the Access Denied Report.We will call the final data frame "AM_missing_plot_df."

# Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,tidyr,vroom,stringr, readr, assertr, ggplot2)
```

Now we can use the original wide data to calculate the missing percentages for different demographic groups and then tack those calculations on to the original data frame "state_missing_totals_df" that we created in the 27-youth-missing-graphics-calculations RMD file.

### AIAN

We will start by finding percent missingness in Native American populations

```{r}
# Get percent of AM students in GT in each state
AM_state_totals_perc <- ocr_17_wide %>%
  group_by(LEA_STATE) %>%
  summarise(tot_AM_enr = (sum(SCH_ENR_AM_M,na.rm = TRUE) + sum(SCH_ENR_AM_F,na.rm = TRUE)), tot_AM_gt_enr = (sum(SCH_GTENR_AM_M,na.rm = TRUE) + sum(SCH_GTENR_AM_F,na.rm = TRUE)), state_AM_perc_gt = (sum(SCH_GTENR_AM_M,na.rm = TRUE) + sum(SCH_GTENR_AM_F,na.rm = TRUE))/(sum(SCH_ENR_AM_M,na.rm = TRUE) + sum(SCH_ENR_AM_F,na.rm = TRUE))) %>%
  ungroup()
```

```{r}
# Add percent of AM students in GT in each state to main DF
AM_missing_df <- state_missing_totals_df %>%
  left_join(AM_state_totals_perc, by = c("LEA_STATE" = "LEA_STATE"))
```

Now we need to break down to populations by students who are in schools that offer GT programs and schools that do not offer GT programs

```{r}
# Find the total number of students in schools that do not offer GT services
AM_enrollment_no_gt_df <- ocr_17_wide %>%
  filter(SCH_GT_IND == "No") %>%
  group_by(LEA_STATE) %>%
  summarise(tot_AM_enr_no_gt = (sum(SCH_ENR_AM_M,na.rm = TRUE) + sum(SCH_ENR_AM_F,na.rm = TRUE))) %>%
  ungroup()
  
```

```{r}
# Find the total number of students in schools that offer GT services
AM_enrollment_w_gt_df <- ocr_17_wide %>%
  filter(SCH_GT_IND == "Yes") %>%
  group_by(LEA_STATE) %>%
  summarise(tot_AM_enr_w_gt = sum(SCH_ENR_AM_M,na.rm = TRUE) + sum(SCH_ENR_AM_F,na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# Add number of students from schools with different service types to full DF
AM_missing_df <- AM_missing_df %>%
  left_join(AM_enrollment_no_gt_df, by = c("LEA_STATE" = "LEA_STATE")) %>%
  left_join(AM_enrollment_w_gt_df, by = c("LEA_STATE" = "LEA_STATE"))

```

Now our data frame has all the columns we need to calculate the number of missing AM students from gt programs.

```{r}
#Find number of students missing from schools that do not offer gifted education
AM_missing_df <- AM_missing_df %>%
  mutate(AM_missing_no_GT_lower = round(tot_AM_enr_no_gt * state_perc_gt), AM_missing_no_GT_upper = round(tot_AM_enr_no_gt * state_nonT1_perc_gt))
```

```{r}
#Find number of underrepresented students from gifted education
AM_missing_df <- AM_missing_df %>%
  mutate(AM_missing_underrep_lower = (round((tot_AM_enr_w_gt * state_perc_gt)) - tot_AM_gt_enr), AM_missing_underrep_upper = (round(tot_AM_enr_w_gt * state_nonT1_perc_gt) - tot_AM_gt_enr))
```

Some under representation values become negative if there is a higher percentage of students in gifted education from that demographic than the percentage of gifted education in the state. We do not consider overrepresentation to be a problem so we will change the negative under representation values to zeros. 

```{r}
#Change negative under representation values to zeros
 AM_missing_df[,15:16][AM_missing_df[,15:16] < 0] <- 0
```

Now we can calculate the upper and lower bounds of the missingness percentages. 

```{r}
# calculate the upper and lower bounds of the missingness percentages.
AM_missing_df <- AM_missing_df %>%
  mutate(AM_perc_missing_lower = ((AM_missing_no_GT_lower + AM_missing_underrep_lower)/(tot_AM_gt_enr + AM_missing_no_GT_lower + AM_missing_underrep_lower)), AM_perc_missing_upper = ((AM_missing_no_GT_upper + AM_missing_underrep_upper)/(tot_AM_gt_enr + AM_missing_no_GT_upper + AM_missing_underrep_upper)))
```

Now we have all the missing percentages calculated we need to do a bit of cleaning before we can plot this data

```{r}
# Select only the columns we want to plot infromation from
AM_plot_fix <- AM_missing_df %>%
  select(LEA_STATE, AM_perc_missing_lower, AM_perc_missing_upper)
```

```{r}
# Find states where lower bound is greater than the upper bound
AM_plot_fix %>%
  filter(AM_perc_missing_lower > AM_perc_missing_upper)
```

In 5 states the lower bound is greater than the upper bound because there is less gifted representation in non title one schools. These states are AZ, DE, FL, MA, MN, and MS. In these states we will flip the upper and lower bound to plot the bars and and I will make a note of this in the report. 

```{r}
#Remove NA values from Alabama, a state that does not report on title1 status which resulted in NA values in the calculations
#Because Alabama does not report on Title1 statues we used the overall percentage for both values for 2015 only
#AM_plot_fix[2,3] <- AM_plot_fix[2,2]

#Change missingness percentage to 100 in states that have no gifted representation at all
AM_plot_fix[8,2:3] <- 1
AM_plot_fix[40,2:3] <- 1
AM_plot_fix[48,2:3] <- 1
```

```{r}
# Add the upperboundry and the lower boundry to the data frame and make sure the missing percentage mins and maxs are the upper and lower boundries
AM_plot_fix <- AM_plot_fix %>%
  mutate(upperbound_add_on = abs(AM_perc_missing_upper - AM_perc_missing_lower))%>%
  rowwise()%>%
  mutate(bottom_bar = min(AM_perc_missing_upper, AM_perc_missing_lower), top_bar = max(AM_perc_missing_upper, AM_perc_missing_lower))
```

```{r}
# Create a df with only the values we need to plot the lower bounds
AM_lower_plot_df <- AM_plot_fix %>%
  select(LEA_STATE, bottom_bar)

AM_lower_plot_df$bound <- "Lower Bound"

colnames(AM_lower_plot_df) <- c("LEA_STATE", "percent_missing", "bound")
```

```{r}
# Create a df with only the values needed to plot the upper bounds
AM_upper_plot_df <- AM_plot_fix %>%
  select(LEA_STATE, upperbound_add_on)

AM_upper_plot_df$bound <- "Upper Bound"

colnames(AM_upper_plot_df) <- c("LEA_STATE", "percent_missing", "bound")
```

```{r}
# Stack upper and lower data frames on top of each other to get our final data frame for plotting. 
AM_missing_plot_df <- rbind(AM_upper_plot_df, AM_lower_plot_df)
```

We created a data frame with just the lower and the upper boundry for percent missing in each state. This will make it easier for use to plot the data in ggplot in a way that matches the graphics from the access denied report. Our data frame is ready and we can now plot it in the rmd file 26-youth-missing-graphics.


