---
title: "27-White-youth-missing-graphics-calculations"
output: html_notebook
---

The purpose of this notebook is to create a data frame that we can use to plot the White youth missing graphic from Figures 16-22 on Pages 115-118 of the Access Denied Report.We will call the final data frame "WH_missing_plot_df."

# Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,tidyr,vroom,stringr, readr, assertr, ggplot2)
```

Now we can use the original wide data to calculate the missing percentages for different demographic groups and then tack those calculations on to the original data frame "state_missing_totals_df" that we created in the 27-youth-missing-graphics-calculations RMD file.

### White

We will use the following code to find percent missingness in White students

```{r}
# Get percent of students in GT in each state
WH_state_totals_perc <- ocr_17_wide %>%
  group_by(LEA_STATE) %>%
  summarise(tot_WH_enr = (sum(SCH_ENR_WH_M,na.rm = TRUE) + sum(SCH_ENR_WH_F,na.rm = TRUE)), tot_WH_gt_enr = (sum(SCH_GTENR_WH_M,na.rm = TRUE) + sum(SCH_GTENR_WH_F,na.rm = TRUE)), state_WH_perc_gt = (sum(SCH_GTENR_WH_M,na.rm = TRUE) + sum(SCH_GTENR_WH_F,na.rm = TRUE))/(sum(SCH_ENR_WH_M,na.rm = TRUE) + sum(SCH_ENR_WH_F,na.rm = TRUE))) %>%
  ungroup()
```

```{r}
# Add percent of students in GT in each state to main DF
WH_missing_df <- state_missing_totals_df %>%
  left_join(WH_state_totals_perc, by = c("LEA_STATE" = "LEA_STATE"))
```

Now we need to break down to populations by students who are in schools that offer GT programs and schools that do not offer GT programs

```{r}
# Find the total number of students in schools that do not offer GT services
WH_enrollment_no_gt_df <- ocr_17_wide %>%
  filter(SCH_GT_IND == "No") %>%
  group_by(LEA_STATE) %>%
  summarise(tot_WH_enr_no_gt = (sum(SCH_ENR_WH_M,na.rm = TRUE) + sum(SCH_ENR_WH_F,na.rm = TRUE))) %>%
  ungroup()
  
```

```{r}
# Find the total number of students in schools that offer GT services
WH_enrollment_w_gt_df <- ocr_17_wide %>%
  filter(SCH_GT_IND == "Yes") %>%
  group_by(LEA_STATE) %>%
  summarise(tot_WH_enr_w_gt = sum(SCH_ENR_WH_M,na.rm = TRUE) + sum(SCH_ENR_WH_F,na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# Add number of students from schools with different service types to full DF
WH_missing_df <- WH_missing_df %>%
  left_join(WH_enrollment_no_gt_df, by = c("LEA_STATE" = "LEA_STATE")) %>%
  left_join(WH_enrollment_w_gt_df, by = c("LEA_STATE" = "LEA_STATE"))

```

Now our data frame has all the columns we need to calculate the number of missing AM students from gt programs.

```{r}
#Find number of students missing from schools that do not offer gifted education
WH_missing_df <- WH_missing_df %>%
  mutate(WH_missing_no_GT_lower = round(tot_WH_enr_no_gt * state_perc_gt), WH_missing_no_GT_upper = round(tot_WH_enr_no_gt * state_nonT1_perc_gt))
```

```{r}
#Find number of underrepresented students from gifted education
WH_missing_df <- WH_missing_df %>%
  mutate(WH_missing_underrep_lower = (round((tot_WH_enr_w_gt * state_perc_gt)) - tot_WH_gt_enr), WH_missing_underrep_upper = (round(tot_WH_enr_w_gt * state_nonT1_perc_gt) - tot_WH_gt_enr))
```

Some under representation values become negative if there is a higher percentage of students in gifted education from that demographic than the percentage of gifted education in the state. We do not consider overrepresentation to be a problem so we will change the negative under representation values to zeros. 

```{r}
#Change negative under representation values to zeros
 WH_missing_df[,15:16][WH_missing_df[,15:16] < 0] <- 0
```

Now we can calculate the upper and lower bounds of the missingness percentages. 

```{r}
# calculate the upper and lower bounds of the missingness percentages.
WH_missing_df <- WH_missing_df %>%
  mutate(WH_perc_missing_lower = ((WH_missing_no_GT_lower + WH_missing_underrep_lower)/(tot_WH_gt_enr + WH_missing_no_GT_lower + WH_missing_underrep_lower)), WH_perc_missing_upper = ((WH_missing_no_GT_upper + WH_missing_underrep_upper)/(tot_WH_gt_enr + WH_missing_no_GT_upper + WH_missing_underrep_upper)))
```

Now we have all the missing percentages calculated we need to do a bit of cleaning before we can plot this data

```{r}
# Select only the columns we want to plot infromation from
WH_plot_fix <- WH_missing_df %>%
  select(LEA_STATE, WH_perc_missing_lower, WH_perc_missing_upper)
```

```{r}
# Find states where lower bound is greater than the upper bound
WH_plot_fix %>%
  filter(WH_perc_missing_lower > WH_perc_missing_upper)
```

In these states we will flip the upper and lower bound to plot the bars and and I will make a note of this in the report. 

```{r}
#Remove NA values from Alabama, a state that does not report on title1 status which resulted in NA values in the calculations
#Because Alabama does not report on Title1 statues we used the overall percentage for both values
#WH_plot_fix[2,3] <- WH_plot_fix[2,2]

#Change missingness percentage to 100 in states that have no gifted representation at all
WH_plot_fix[8,2:3] <- 1
WH_plot_fix[40,2:3] <- 1
```

```{r}
# Add the upperboundry and the lower boundry to the data frame and make sure the missing percentage mins and maxs are the upper and lower boundries
WH_plot_fix <- WH_plot_fix %>%
  mutate(upperbound_add_on = abs(WH_perc_missing_upper - WH_perc_missing_lower))%>%
  rowwise()%>%
  mutate(bottom_bar = min(WH_perc_missing_upper, WH_perc_missing_lower), top_bar = max(WH_perc_missing_upper, WH_perc_missing_lower))
```

```{r}
# Create a df with only the values we need to plot the lower bounds
WH_lower_plot_df <- WH_plot_fix %>%
  select(LEA_STATE, bottom_bar)

WH_lower_plot_df$bound <- "Lower Bound"

colnames(WH_lower_plot_df) <- c("LEA_STATE", "percent_missing", "bound")
```

```{r}
# Create a df with only the values needed to plot the upper bounds
WH_upper_plot_df <- WH_plot_fix %>%
  select(LEA_STATE, upperbound_add_on)

WH_upper_plot_df$bound <- "Upper Bound"

colnames(WH_upper_plot_df) <- c("LEA_STATE", "percent_missing", "bound")
```

```{r}
# Stack upper and lower data frames on top of each other to get our final data frame for plotting. 
WH_missing_plot_df <- rbind(WH_upper_plot_df, WH_lower_plot_df)
```

We created a data frame with just the lower and the upper boundary for percent missing in each state. This will make it easier for use to plot the data in ggplot in a way that matches the graphics from the access denied report. Our data frame is ready and we can now plot it in the rmd file 26-youth-missing-graphics.


