---
title: "27-MEGATABLE-youth-missing-graphics-calculations"
output: html_notebook
---

The purpose of this notebook is to create a data frame that we can use to plot the youth missing graphics for all demographics in the Shiny app from Figures 16-22 on Pages 115-118 of the Access Denied Report.We will call the final data frame "full_missing_plots_df"

# Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,tidyr,vroom,stringr, readr, assertr, ggplot2)
```

# Create 2015 df for AIAN

```{r}
# Add total missing percentage for plotly to print out
AM_2015 <- AM_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
AM_2015$year <- 2015
AM_2015$group <- "AIAN"
```

# Create 2015 df for Asian

```{r}
# Add total missing percentage for plotly to print out
AS_2015 <- AS_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
AS_2015$year <- 2015
AS_2015$group <- "Asian"
```

# Create 2015 df for Black

```{r}
# Add total missing percentage for plotly to print out
BL_2015 <- BL_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
BL_2015$year <- 2015
BL_2015$group <- "Black"
```

# Create 2015 df for Latinx

```{r}
# Add total missing percentage for plotly to print out
HI_2015 <- HI_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
HI_2015$year <- 2015
HI_2015$group <- "Latinx"
```

# Create 2015 df for NHPI

```{r}
# Add total missing percentage for plotly to print out
HP_2015 <- HP_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
HP_2015$year <- 2015
HP_2015$group <- "NHPI"
```

# Create 2015 df for TMR

```{r}
# Add total missing percentage for plotly to print out
TR_2015 <- TR_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
TR_2015$year <- 2015
TR_2015$group <- "TMR"
```

# Create 2015 df for White

```{r}
# Add total missing percentage for plotly to print out
WH_2015 <- WH_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
WH_2015$year <- 2015
WH_2015$group <- "White"
```


# Bind all groups together to get the 2015 data frame

```{r}
# Stack upper and lower data frames on top of each other to get our final data frame for plotting. 
full_plot_df_2015 <- rbind(AM_2015, AS_2015, BL_2015, HI_2015, HP_2015, TR_2015, WH_2015)
```

# Save the full data frame to CSV for easy plotting

```{r}
write_csv(full_plot_df_2015, "C:/Users/donse/OneDrive/Documents/DSSG/gateways/full_plot_df_2015.csv")
```


#### Create 2017 DF------------------------------------------------------------------------------------------------------------

# Create 2017 df for AIAN

```{r}
# Add total missing percentage for plotly to print out
AM_2017 <- AM_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
AM_2017$year <- 2017
AM_2017$group <- "AIAN"
```

# Create 2017 df for Asian

```{r}
# Add total missing percentage for plotly to print out
AS_2017 <- AS_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
AS_2017$year <- 2017
AS_2017$group <- "Asian"
```

# Create 2017 df for Black

```{r}
# Add total missing percentage for plotly to print out
BL_2017 <- BL_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
BL_2017$year <- 2017
BL_2017$group <- "Black"
```

# Create 2017 df for Latinx

```{r}
# Add total missing percentage for plotly to print out
HI_2017 <- HI_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
HI_2017$year <- 2017
HI_2017$group <- "Latinx"
```

# Create 2017 df for NHPI

```{r}
# Add total missing percentage for plotly to print out
HP_2017 <- HP_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
HP_2017$year <- 2017
HP_2017$group <- "NHPI"
```

# Create 2017 df for TMR

```{r}
# Add total missing percentage for plotly to print out
TR_2017 <- TR_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
TR_2017$year <- 2017
TR_2017$group <- "TMR"
```

# Create 2017 df for White

```{r}
# Add total missing percentage for plotly to print out
WH_2017 <- WH_missing_plot_df %>%
  group_by(LEA_STATE) %>%
  mutate(total_missing_perc = sum(percent_missing)) %>%
  ungroup()
```

```{r}
# Add on vairables for year and group
WH_2017$year <- 2017
WH_2017$group <- "White"
```


# Bind all groups together to get the 2015 data frame

```{r}
# Stack upper and lower data frames on top of each other to get our final data frame for plotting. 
full_plot_df_2017 <- rbind(AM_2017, AS_2017, BL_2017, HI_2017, HP_2017, TR_2017, WH_2017)
```

# Combine 2015 and 2017 data frame -----------------------------------------------------------------------------------------

```{r}
#Combine to form final plotting data frame
full_missing_plots_df <- rbind(full_plot_df_2015, full_plot_df_2017)
```

# Save the plot to CSV in box with write_csv


